Memo: MWA beam-correction formulae

### 2013-08-18 articles/Report-MWA-Beam/mwa-beam.tex
Some changes on beam in cal
Fixing sign error in beam\!

\section{Analytical calculation of voltage gain}
The analytical gain of a tile is calculated from the analytical gain of a dipole with ground screen, multiplied by a factor $\eta$ that represents how coherent the dipoles add together. The (complex) tile factor $\eta$ is calculated with:\footnote{Taken from the RTS (2013-08-13), file \texttt{InstrumentalCalibration.c} written by D. Mitchell}
\begin{equation}
 \eta = \frac{1}{16}\sum\limits_{a=1}^{16} e^{\frac{-2 \pi i}{\lambda} \left(\Delta x_a \sin \mathcal{Z} \sin \chi + \Delta y_a \sin \mathcal{Z} \cos \chi + \Delta z_a \cos \mathcal{Z} - \Delta \tau_a \right) },
\end{equation}
where $\lambda$ is the wavelength, $\Delta x_a, \Delta y_a$ and $\Delta z_a$ are the relative positions of dipole $a$ in the tile, $\mathcal{Z}$ is zenith angle, $\chi$ is the azimuth and $\Delta \tau_a$ is the delay applied on dipole $a$ by the beam former. The $2 \times 2$ matrix $\textbf{E}$ describing the electrical gains of an individual dipole\footnote{Also described here: http://mwa-lfd.haystack.mit.edu/twiki/bin/view/Main/PrimaryBeamModels by D. Mitchell} is calculated with:
\begin{equation}
\mathbf{E} = \frac{2 \sin \left( \frac{2 \pi h}{\lambda} \cos \mathcal{Z}\right)}{2 \sin \left( \frac{2 \pi h}{\lambda}\right)}
\left( \begin{array}{ccc}
\cos l \cos \delta + \sin l \sin \delta \cos \textrm{ha} & ; & -\sin l \sin \textrm{ha} \\
\sin \delta \sin \textrm{ha} & ; & \cos \textrm{ha}
\end{array} \right).
\end{equation}
Here, $l$ is latitude, $\delta$ is declanation, $\textrm{ha}$ is hour angle and $h$ is the height of tile. The factors of two cancel out, but are left there because the denominator now forms the normalization towards zenith. Finally, the Jones gain matrix $\textbf{B}$ of the full tile is calculated by multiplying the dipole matrix with the tile factor,
\begin{equation}
\mathbf{B} = \eta \mathbf{E}.
\end{equation}
\section{Applying beam gains in image space}
Given the $2\times 2$ complex Jones matrix $\mathbf{B}$ that represents the voltage gain of the beam in a certain direction, the RTS corrects a polarized pixel, given by the vector $\textbf{v}' = \left(v_\textrm{xx}, \textrm{real}(v_\textrm{xy}), \textrm{imag}(v_\textrm{xy}), v_\textrm{yy} \right)^T$, for that direction in the image plane to a beam-corrected vector containing the Stokes values with the following calculation:\footnote{Taken from the RTS (2013-08-13), file \texttt{pb\_correct.c} written by S. Ord}
\begin{equation} \label{apply-beam-image-space}
 \textbf{s} =
\mathbf{T} \left[ \left(\mathbf{B} \otimes \mathbf{B}^{*T} \right)^* \left(\mathbf{B} \otimes \mathbf{B}^{*T} \right)\right]^{-1} \mathbf{Y} \mathbf{v}',
\end{equation}
\textbf{Note:} the formula above is how the RTS seems to do it, but it may not be correct. Mitch is going to check this, as it should be:
\begin{equation}
 \textbf{s} =
\mathbf{T} \left(\mathbf{B} \otimes \mathbf{B}^{*T} \right)^{-1} \mathbf{Y} \mathbf{v}',
\end{equation}
with
\begin{equation}
\textbf{s} = \left( \begin{array}{c}
s_\textrm{\tiny{I}} \\
s_\textrm{\tiny{Q}} \\
s_\textrm{\tiny{U}} \\
s_\textrm{\tiny{V}} \end{array} \right),
T = \left( \begin{array}{cccc}
1 & 0 & 0 & 1 \\
1 & 0 & 0 & -1 \\
0 & 1 & 1 & 0 \\
0 & -i & -i & 0 \end{array} \right), \mathbf{Y} = \left( \begin{array}{cccc}
1 & 0 & 0 & 0 \\
0 & 1 & i & 0 \\
0 & 1 & -i & 0 \\
0 & 0 & 0 & 1
\end{array} \right),
\end{equation}
where $\otimes$ the Kronecker product, $\mathbf{A}^*$ the conjugate transpose of matrix $\mathbf{A}$ and $\mathbf{A}^{*T}$ the matrix $\mathbf{A}$ where each element has been conjugated ($A_{ij}$ = $\overline{A^{*T}_{ij}}$). The output $\textbf{s}$ is a vector with the Stokes parameters. The matrix $\mathbf{T}$ converts from instrumental basis (xx, xy, yx, yy) to the Stokes parameter and hence is applied last.

>>>>

\section{Analytical calculation of voltage gain}
The analytical gain of a tile is calculated from the analytical gain of a dipole with ground screen, multiplied by a factor $\eta$ that represents how coherent the dipoles add together. The (complex) tile factor $\eta$ is calculated with:\footnote{Taken from the RTS (2013-08-13), file \texttt{InstrumentalCalibration.c} written by D. Mitchell}
\begin{equation}
 \eta = \frac{1}{16}\sum\limits_{a=1}^{16} e^{\frac{-2 \pi i}{\lambda} \left(\Delta x_a \sin \mathcal{Z} \sin \chi + \Delta y_a \sin \mathcal{Z} \cos \chi + \Delta z_a \cos \mathcal{Z} - \Delta \tau_a \right) },
\end{equation}
where $\lambda$ is the wavelength, $\Delta x_a, \Delta y_a$ and $\Delta z_a$ are the relative positions of dipole $a$ in the tile, $\mathcal{Z}$ is zenith angle, $\chi$ is the azimuth and $\Delta \tau_a$ is the delay applied on dipole $a$ by the beam former. Antennae are 1.1 \textit{m} separated from each other within tiles and currently are all assumed to have a differential height $\Delta z_i$ of 0. The MWA tools pb scripts assume the antennae are 0.278~\textit{m} tall while the RTS assumes 0.3~\textit{m}. The $2 \times 2$ matrix $\textbf{E}$ describing the electrical gains of an individual dipole\footnote{Also described here: http://mwa-lfd.haystack.mit.edu/twiki/bin/view/Main/PrimaryBeamModels by D. Mitchell} is calculated with:
\begin{equation}
\mathbf{E} = \frac{2 \sin \left( \frac{2 \pi h}{\lambda} \cos \mathcal{Z}\right)}{2 \sin \left( \frac{2 \pi h}{\lambda}\right)}
\left( \begin{array}{ccc}
\cos l \cos \delta + \sin l \sin \delta \cos \textrm{ha} & ; & -\sin l \sin \textrm{ha} \\
\sin \delta \sin \textrm{ha} & ; & \cos \textrm{ha}
\end{array} \right).
\end{equation}
Here, $l$ is latitude, $\delta$ is declanation, $\textrm{ha}$ is hour angle and $h$ is the height of tile. The factors of two cancel out, but are left there because the denominator now forms the normalization towards zenith. Finally, the Jones gain matrix $\textbf{B}$ of the full tile is calculated by multiplying the dipole matrix with the tile factor,
\begin{equation}
\mathbf{B} = \eta \mathbf{E}.
\end{equation}
\section{Applying beam gains in image space}
Given the $2\times 2$ complex Jones matrix $\mathbf{B}$ that represents the voltage gain of the beam in a certain direction, the RTS corrects a polarized pixel, given by the vector $\textbf{v}' = \left(v_\textrm{xx}, \textrm{real}(v_\textrm{xy}), \textrm{imag}(v_\textrm{xy}), v_\textrm{yy} \right)^T$, for that direction in the image plane to a beam-corrected vector containing the Stokes values with the following calculation:\footnote{Taken from the RTS (2013-08-13), file \texttt{pb\_correct.c} written by S. Ord}
\begin{equation} \label{apply-beam-image-space}
 \textbf{s} =
\mathbf{T} \left[ \left(\mathbf{B} \otimes \mathbf{B}^{*T} \right)^* \left(\mathbf{B} \otimes \mathbf{B}^{*T} \right)\right]^{-1} \mathbf{Y} \mathbf{v}',
\end{equation}
\textbf{Note:} the formula above is how the RTS seems to do it, but it may not be correct. Mitch is going to check this, as it should be:
\begin{equation}
 \textbf{s} =
\mathbf{T} \left(\mathbf{B} \otimes \mathbf{B}^{*T} \right)^{-1} \mathbf{Y} \mathbf{v}',
\end{equation}
with
\begin{equation}
\textbf{s} = \left( \begin{array}{c}
s_\textrm{\tiny{I}} \\
s_\textrm{\tiny{Q}} \\
s_\textrm{\tiny{U}} \\
s_\textrm{\tiny{V}} \end{array} \right),
T = \left( \begin{array}{cccc}
1 & 0 & 0 & 1 \\
1 & 0 & 0 & -1 \\
0 & 1 & 1 & 0 \\
0 & -i & -i & 0 \end{array} \right), \mathbf{Y} = \left( \begin{array}{cccc}
1 & 0 & 0 & 0 \\
0 & 1 & i & 0 \\
0 & 1 & -i & 0 \\
0 & 0 & 0 & 1
\end{array} \right),
\end{equation}
where $\otimes$ the Kronecker product, $\mathbf{A}^*$ the conjugate transpose of matrix $\mathbf{A}$ and $\mathbf{A}^{*T}$ the matrix $\mathbf{A}$ where each element has been conjugated ($A_{ij}$ = $\overline{A^{*T}_{ij}}$). The output $\textbf{s}$ is a vector with the Stokes parameters. The matrix $\mathbf{T}$ converts from instrumental basis (xx, xy, yx, yy) to the Stokes parameter and hence is applied last.
